---
title: "STAT545_Gapminder_Updated"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Updated Gapminder Homework Assignment

**Note:** A previous file shows a data analysis I attempted using a dataset that I created myself. I misunderstood the homework instructions, so I am performing the correct data analysis here. In the original file [STAT545_gapminder_hw01](https://github.com/STAT545-UBC-students/hw01-JasmineLib/blob/master/STAT545_Gapminder_hw01.md) I performed a t-test on two of my own data sets, and studied the resulting p-values (I included links on how I learned to do this in the document), which may be useful to some of you reading this now!


###Explore the iris dataframe

```{r}
library(tidyverse)
data(iris)
head(iris)
ncol(iris)
nrow(iris)
summary(iris)

```

There are 5 columns and 150 rows in this dataset.

### Explore variables in the iris dataframe.
2. In particular, I want to observe the species in this dataframe.
```{r}
iris$Species
```
I conclude there are 3 different species (by looking at the outpud "levels"): setosa, versicolor and virginica



3. Compare sepal length (Sepal.Length) of setosa and versicolor species
To do this, I used the [cm005 notes on piping](http://stat545.com/Classroom/notes/cm005.nb.html) as well as a tutorial on finding the [mean of a column in r](https://stackoverflow.com/questions/23163863/mean-of-a-column-in-a-data-frame-given-the-columns-name)
```{r}
library(tidyverse)
setosa = iris %>% 

  filter(Species == "setosa") %>% 
  select(Sepal.Length)
  

versicolor = iris %>% 
  filter(Species == "versicolor") %>% 
  select(Sepal.Length)

# I can check the mean of the Sepal length for setosa and versicolor species using the mean function:
mean(setosa[["Sepal.Length"]])
mean(versicolor[["Sepal.Length"]])

```

4. I can calculate the variance of the sepal length in the setosa species using the var function:
```{r}
var(setosa)
```
5. I can check this result manually, by calculating the variance. The equation for variance can be found [here](http://www.r-tutor.com/elementary-statistics/numerical-measures/variance)


breaking down the equation:
  1.find the differences between each entry in setosa compared to the mean:
  (setosa$Sepal.Length - 5.006)
  2. Take the square of resulting vector:
(setosa$Sepal.Length - 5.006)^2
3. Take the sum of the squared vector:
sum((setosa$Sepal.Length - 5.006)^2)
4. Divide by the number of rows in setosa -1
sum((setosa$Sepal.Length - 5.006)^2)/(nrow(setosa)-1)

```{r}

#overall, the equation comes to:
sum((setosa$Sepal.Length-5.006)^2)/(nrow(setosa)-1)



```



 

1. Load the gapminder package and explore the data inside it using the head() function
```{r}
library(gapminder)
head(gapminder)
```

2. Load tidyverse and explore the average life expectancy in 1952 in the continent with the most entries.
```{r}
library(tidyverse)
gapminder %>% 
  select(continent, lifeExp, year) %>% 
  filter(year =="1952") %>% 
  summary() #this outputs the 5 continents, so I can ensure there are values in each one, and there are.



```
3. From above, I can see that there are the most entries in the continent of Africa, so I will explore Asia's average life expectancy in 1952.
```{r}
gapminder %>% 
  select(continent, lifeExp, year) %>% 
  filter(continent== "Asia" & year =="1952") %>% 
  summary()

#using the summary function, I determine that the mean life expectancy in Africa in 1952 is 46.31 years. 
  
```

