---
title: "STAT 545 Gapminder Homework"
output: github_document
---
##Goal: Compare average fluorescence intensity of two Mutant reporter proteins to that of Wildtype reporter

### Create datasets:

Create three vectors containing values of fluorescence intensity values from independent experiments for both mutants and the wildtype
```{r}
 
fluor_m1 = c(55.74, 70.81, 36.74, 65.1, 57.22, 77.69, 53.58, 63.94, 50.8, 64.05, 45.09, 58.86)
fluor_m2 = c(43.37, 35.23, 67.87, 53.3, 63.4, 56.01, 66.81, 58.51, 67.98, 53.34, 66.42, 45.71)
fluor_wt = c(12.23, 10.72, 17.28, 16.55, 22.25, 15.81, 17.32, 12.8, 16.5, 14.71, 19.71, 17.86)
```

### Compute the mean of each dataset: 
Compute the average of the values in each vector using the 'mean' function
Note: to remove missing values, use na.rm (see ?mean in search.)

```{r}
(mean_fluorm1 = mean(fluor_m1))
(mean_fluorm2 = mean(fluor_m2))
(mean_fluorwt = mean(fluor_wt))

```
Conclusion: The means of the mutants appear to be different from that of the wildtype. Will perform a t-test using the t-test function in R. 

### Perform a t-test
A useful link to find out what t-test to use: https://researchbasics.education.uconn.edu/t-test/

####Compare fluorescence of mutant 1 to that of wildtype:
```{r}
(fluor_m1wt_ttest = t.test(fluor_m1, fluor_wt))
```

Can save and print the p-value from a t-test as a variable: 
Useful link for learning to do this: https://stackoverflow.com/questions/31205554/output-p-value-from-a-t-test-in-r
```{r}
(m1wt_ttest_pval = fluor_m1wt_ttest$p.value)
(type(m1wt_ttest_pval))
```
Is the p-value below 0.05? Visually yes, but I can check this using an if/else statement
Useful link for if-else statements: https://stackoverflow.com/questions/25885358/if-else-if-else-statement-and-brackets
```{r}
 if (int(m1wt_ttest_pval) < 0.05) { 
  print("Accept alternate hypothesis")
   } else if (int(m1wt_ttest_pval) >= 0.05) {
  print("Accept null hypothesis")
     else {
       print("error")
     }
} 

