---
title: "STAT545_Gapminder_Updated"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Updated Gapminder Homework Assignment

**Note:** A previous file shows a data analysis I attempted using a dataset that I created myself. I misunderstood the homework instructions, so I am performing the correct data analysis here. In the original file [STAT545_gapminder_hw01](https://github.com/STAT545-UBC-students/hw01-JasmineLib/blob/master/STAT545_Gapminder_hw01.md) I performed a t-test on two of my data sets, and studied the resulting p-values (with links on how I learned to do this), which may be useful to some of you reading this now!

Part 1
###Explore the iris dataframe

```{r}
data(iris)
head(iris)
ncol(iris)
nrow(iris)
summary(iris)

```

There are 5 columns and 150 rows in this dataset.

### Explore variables in the iris dataframe.
2. In particular, I want to observe the species in this dataframe.
```{r}
iris$Species
```
I conclude there are 3 different species (by looking at the outpud "levels"): setosa, versicolor and virginica



3. Compare sepal length (Sepal.Length) of setosa and versicolor species
```{r}

```




 

1. Load the gapminder package and explore the data inside it using the head() function
```{r}
library(gapminder)
head(gapminder)
```

2. Load tidyverse and explore the average life expectancy in 1952 in the continent with the most entries.
```{r}
library(tidyverse)
gapminder %>% 
  select(continent, lifeExp, year) %>% 
  filter(year =="1952") %>% 
  summary() #this outputs the 5 continents, so I can ensure there are values in each one, and there are.



```
3. From above, I can see that there are the most entries in the continent of Africa, so I will explore Asia's average life expectancy in 1952.
```{r}
gapminder %>% 
  select(continent, lifeExp, year) %>% 
  filter(continent== "Asia" & year =="1952") %>% 
  summary()

#using the summary function, I determine that the mean life expectancy in Africa in 1952 is 46.31 years. 
  
```

